I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-19 Paper/introduction.tex
introduction

While the verification engines are generic, the benchmarks usually focus on local robustness, i.e. given a Network, an image and a small neighbourhood around this image, 
is it the case that all the images in the neighbourhood are classified in the same way.
While some quite large DNNs (e.g. ResNet with tens of thousands of neurons) can be verified very efficiently (tens of seconds per input to check local robustness) \cite{crown}, with all inputs either certified robust or an attack on robustness is found in the neighbourhood;
some smaller DNNs (with hundreds of neurons, only using the simpler ReLU activation function) cannot be analysed fully, with $12-20 \%$ of inputs where neither of the decisions can be reached \cite{crown} and Table \ref{tab:example}. The main difference between these DNNs with very different behaviours lies in that the DNNs which are trained to be robust are much easier to verify, while the DNNs trained in a "natural" way are much harder to verify.

In this paper, we focus on DNNs trained in a "natural", 
%uncovering what makes the DNNs trained in a natural way so hard to verify (
because for "easier" DNNs, adequate methods already exist. 
To do so, we analyse the abstraction mechanisms at the heart of several efficient algorithms, namely Eran-DeepPoly \cite{deeppoly}, the Linear Programming approximation \cite{MILP}, PRIMA \cite{prima}, and different versions of ($\alpha$)($\beta$)-CROWN \cite{crown}. All these algorithms compute lower or/and upper bounds for the values of neurons (abstraction on values) for inputs in the considered input region, and conclude based on such bounds. For instance, if for all image $I'$ in the neighbourhood of image $I$, we have $weight_{I'}(n'-n) < 0$ for $n$ the output neuron corresponding to the expected class, then we know that the DNN is robust in the neighbourhood of image $I$. We restrict the formal study to DNNs using only the standard ReLU activation function, although nothing specific prevents the results to be extended to more general architectures. We uncover that {\em compensations} (see next paragraph) is the phenomenon creating inaccuracies. We verified experimentally that indeed DNNs trained in a natural way have much more heavy compensating pairs than DNNs trained in a robust way.

@@

On such complex DNNs, PRIMA and $\alpha$-$\beta$ CROWN resort to a "refined" path, where the bounds {\em on the first few layers} are refined \cite{MILP2} using an exact MILP encoding. In our algorithm, we do not use an exact encoding but a partial one with the most important ReLU nodes obtained by considering the compensation strength. As it is more efficient, this can be pushed to all layers. This would be infeasible without the selection based on the compensation. The refined version of $\alpha$-$\beta$ CROWN is particularly accurate on small DNNs. As the depth grows, the more work is left to BaB and our algorithm is more accurate, lowering the gap of verified images to the upper bound down from $17.6\%$ to $12\%$.

>>>>

While the verification engines are generic, the benchmarks usually focus on local robustness, i.e. given a Network, an image and a small neighbourhood around this image, 
is it the case that all the images in the neighbourhood are classified in the same way.
While some quite large DNNs (e.g. ResNet with tens of thousands of neurons) can be verified very efficiently (tens of seconds per input) \cite{crown}, with all inputs either certified robust or an attack on robustness is found; some smaller DNNs (with hundreds of neurons, only using the simpler ReLU activation function) cannot be analysed fully, with $12-20\%$ of inputs where neither of the decisions can be reached \cite{crown} and Table \ref{tab:example}. The main difference between these DNNs with very different behaviours lies in that the DNNs which are trained to be robust are much easier to verify, while the DNNs trained in a "natural" way are much harder to verify.

In this paper, we focus on DNNs trained in a "natural", 
%uncovering what makes the DNNs trained in a natural way so hard to verify (
because for "easier" DNNs, adequate methods already exist. 
To do so, we analyse the abstraction mechanisms at the heart of several efficient algorithms, namely Eran-DeepPoly \cite{deeppoly}, the Linear Programming approximation \cite{MILP}, PRIMA \cite{prima}, and different versions of ($\alpha$)($\beta$)-CROWN \cite{crown}. All these algorithms compute lower or/and upper bounds for the values of neurons (abstraction on values) for inputs in the considered input region, and conclude based on such bounds. For instance, if for all image $I'$ in the neighbourhood of image $I$, we have $weight_{I'}(n'-n) < 0$ for $n$ the output neuron corresponding to the expected class, then we know that the DNN is robust in the neighbourhood of image $I$. We restrict the formal study to DNNs using only the standard ReLU activation function, although nothing specific prevents the results to be extended to more general architectures. We uncover that {\em compensations} 
(see next paragraph) is the phenomenon creating inaccuracies. We verified experimentally that DNNs trained in a natural way have much more heavy compensating pairs than DNNs trained in a robust way.

@@

On such complex DNNs, PRIMA and $\alpha$-$\beta$ CROWN resort to a "refined" path, where the bounds {\em on the first few layers} are refined \cite{MILP2} using an exact MILP encoding. In our algorithm, we do not use an exact encoding but a partial one with the most important ReLU nodes obtained by considering the compensation strength. As it is more efficient, this can be pushed to all layers. This would be infeasible without the selection based on the compensation. The refined version of $\alpha$-$\beta$ CROWN is particularly accurate on small DNNs. As the depth grows, the more work is left to BaB and our algorithm is more accurate, lowering the gap of verified images to the upper bound down from 
$20\%$ to $16.2\%$ and from $17.6\%$ to $12\%$ (depth 9, Table \ref{tab:example}). On larger naturally learnt DNNs (3000 neurons), BaB only verifies $18.5\%$ more images than DeepPoly (PRIMA only $6.5\%$), while Compensate verifies $39\%$ more images than DeepPoly (Table \ref{tab:example3}).
